Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-09-01 paper/paper.tex
complete new version updated referee reply

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
\sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
{\bf We argue that the results from the multiphase model 
provides a correct interpretation of the data.
In that case the clump velocity dispersion implies a dynamical mass of
$2\times 10^{9}$ M$_{\odot}$,  ten times its baryonic mass.}
If future kinematic maps of \tol\ confirm the velocities suggested by
the multiphase model, it would provide additional support to expect
such kinematic state in primordial \lya emitting galaxies {\bf which in
turn can help to constrain star formation and feedback in the first
generation of galaxies.}
\end{abstract}

@@

{\bf Furthermore, observations of other Compact Dwarf Galaxies (CDGs) also show 
that the kinematics range from pure rotation and low velocity
dispersion to high velocity dispersion without rotaion
\citep{2015A&A...577A..21C,2017A&A...600A.125C}. }

@@

The observed flux for the \lya line is $\sim
8.1\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \citep{Thuan97}.
The \lya Equivalent Width is $70$\AA\ and its H$\beta$ flux is 
$1.62\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$
which gives a \lya/H$\beta$ flux ratio of
4.9$\pm$0.1 \citep{Izotov04}.
Comparing the \lya/H$\beta$ ratio with the theoretical
expectation from case B recombination of $23.3$ \citep{Hummer1987} one
can estimate an escape fraction of $20$\% for \lya radiation.
Figure \ref{fig:results} shows \tol's \lya\ profile reported by
\cite{mashesse03}. This measurement was made with the Space Telescope
Imaging Spectrograph on board the Hubble Space Telescope, these
observations have a spectral resolution of $\sim 37$\kms\ at the
\lya\ wavelength.

@@

In order to map out this large parameter space, we randomly drew
$2500$ sets of parameters within a observationally realistic range,
based on the considerations of \citet{Laursen2013ApJ...766..124L}, 
yielding a large variety of single-, double- and triple-peaked
spectra. 
The full analysis of the spectral features as well as
more details on the radiative transfer are presented by
\citet{Gronke2016}.

@@

We use an analytical approximation that captures the most important
effects of rotation onto the \lya line. 
We defer the reader to \citet{GaravitoCamargo2014} for complete
details on the explicit form of this approximation.
To fully explore the
parameter space using a Markov Chain Monte Carlo (MCMC) calculation with the
\texttt{emcee} Python library \citep{2013PASP..125..306F}. \texttt{emcee} 
is an open source optimized implementation of the affine-invariant 
MCMC sampler \citep{goodman2010ensemble}. 
The algorithm creates a number of walkers that,
during a sufficient number of steps, generate parameters' combinations
for a specific model.
For each timestep, the code calculates the likelihood of the
combination with respect to the observational data.
The walkers explore the parameter space sampling the Gaussian likelihood
function built as $\propto \exp(-\chi^2/2)$, where the $\chi^2$ follows
the definition in Equation \ref{eq:chi2}. 
We do not have a closed analytic expression for $\hat{f}$, we compute it
by numerical integration of the analytical approximations found in
\citet{GaravitoCamargo2014}.

@@

The low p-values are illustrated by the results shown in Figure \ref{multiphaseresults}.
Left column shows the difference between the integrated distributions
of the full sample (2500 input models) and the $1\%$ models with the lowest
$\chi^2/d.o.f.$, where we use the total number of degrees of freedom,
$d.o.f.=104$. 
The right column shows the actual $\chi^2/d.o.f.$ 
and its corresponding parameter value. 
Under these conditions we find $\sigma_{\rm cl}=$\sigmaclump,
$v_{\infty {\rm, cl}}=$\inftyclump\ and $P_{\rm cl}=$\probaclump.

Qualitatively this result can be understood as follows.
Due to the large fraction of \lya photons being emitted within the
moving clumps ($P_{\rm cl}\sim 1$) the `intrinsic' profile follows 
the clump kinematics. In other words, the width of the intrinsic line
is set by $\sigma_{\mathrm{cl}}$ and its median offset mainly by
$v_{\infty,{\rm cl}}$ and $\beta_{\rm cl}$.
Furthermore, the relatively low mean number of clumps per line of
sight, $f_{\mathrm{cl}}$, combined with the high velocity dispersion
of the clumps implies existence of low-density inter-clump 
regions where \lya photons can freely propagate.
This gives result to an emergent spectrum close to the intrinsic one, explaining the high
flux at the line's center.

@@

Following the methodology of \citet{2011ApJ...726..108T} we estimate
that a dark matter halo of $\sim 6\times 10^{11}$ M$_{\odot}$ in
virial mass {\bf and virial radius $220$ kpc } could explain this dynamical mass.
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the median mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P,2016ApJ...832..169P}.  
{\bf This estimate should be considered an upper bound, lower values
  could be achieved if one considers instead a cored profile.}
Figure \ref{fig:mass} shows the enclosed mass as a function of radius
for different dark matter halos together with the \tol's dynamical
mass estimate. 
{\bf This estimate and the validity of its assumptions can only be
  verified by detailed mass modelling once \tol's full rotation curve
  becomes available.}

{\bf New IFU observations are needed to confirm \tol's kinematics.
This could be done with the Multi Unit Spectroscopic Explorer (MUSE) 
\citep{2014Msngr.157...13B}, the Gemini Multi-Object Spectrographs (GMOS) 
\citep{2004PASP..116..425H} or the Fibre Large Array Multi Element
Spectrograph (FLAMES) \citep{2002Msngr.110....1P} as they have the spatial
resolution and field of view required to perform the kind of study
presented by \citet{2015A&A...577A..21C} on CDGs or \citet{Herenz16}
onnearby \lya\ emitting galaxies.}

\begin{figure}
\begin{center}
\includegraphics[width=0.46\textwidth]{enclosed_mass.pdf}
\caption{Enclosed mass in a spherical system as a function of radius.
  Lines correspond to the expectation for dark matter halos following
  an NFW profile with
  different viral masses as shown in the legend, the arrow corresponds
  to the dynamical mass lower bounds for \tol.  
  The dynamical mass estimates for \tol\ are consistent with the
  galaxy being hosted by a dark matter halo of
  $\sim 6\times10^{11}$ M$_{\odot}$ in mass.
  {\bf Lower mass estimates could be possible if one considers instead a
  cored profile. }
    \label{fig:mass}} 
\end{center}
\end{figure}

>>>>

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
\sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
{\bf We argue that the results from the multiphase model 
provide a correct interpretation of the data.
In that case the clump velocity dispersion implies a dynamical mass of
$2\times 10^{9}$ M$_{\odot}$,  ten times its baryonic mass.}
If future kinematic maps of \tol\ confirm the velocities suggested by
the multiphase model, it would provide additional support to expect
such kinematic state in primordial \lya emitting galaxies, opening
{\bf the opportunity to constrain models of star formation and
  feedback in the first generation of galaxies.}
\end{abstract}

@@

{\bf Furthermore, observations of other Compact Dwarf Galaxies (CDGs) also show 
that the kinematics range from pure rotation and low velocity
dispersion to high velocity dispersion without rotation
\citep{2015A&A...577A..21C,2017A&A...600A.125C}. }

@@

The observed flux for the \lya line is $\sim
8.1\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \citep{Thuan97}.
The \lya Equivalent Width is $70$\AA\ and its H$\beta$ flux is 
$1.62\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$
which gives a \lya/H$\beta$ flux ratio of
4.9$\pm$0.1 \citep{Izotov04}.
Comparing the \lya/H$\beta$ ratio with the theoretical
expectation from case B recombination of $23.3$ \citep{Hummer1987} one
can estimate an escape fraction of $20$\% for \lya radiation.
Figure \ref{fig:results} shows \tol's \lya\ profile reported by
\cite{mashesse03}. This measurement was made with the Space Telescope
Imaging Spectrograph on board the Hubble Space Telescope, with a
spectral resolution of $\sim 37$\kms\ at the  \lya\ wavelength.

@@

In order to map out this large parameter space, we randomly drew
$2500$ sets of parameters within an observationally realistic range,
based on the considerations of \citet{Laursen2013ApJ...766..124L}, 
yielding a large variety of single-, double- and triple-peaked
spectra. 
The full analysis of the spectral features as well as
more details on the radiative transfer are presented by
\citet{Gronke2016}.

@@

We use an analytical approximation that captures the most important
effects of rotation onto the \lya line. 
We defer the reader to \citet{GaravitoCamargo2014} for complete
details on the explicit form of this approximation.
To fully explore the parameter space we perform Markov Chain Monte
Carlo (MCMC) calculations with the \texttt{emcee} Python library
\citep{2013PASP..125..306F}. \texttt{emcee}  is an open source
optimized implementation of the affine-invariant  MCMC sampler
\citep{goodman2010ensemble}.  
The algorithm creates a number of walkers that,
during a sufficient number of steps, generate parameters' combinations
for a specific model.
For each timestep, the code calculates the likelihood of the
combination with respect to the observational data.
The walkers explore the parameter space sampling the Gaussian likelihood
function built as $\propto \exp(-\chi^2/2)$, where the $\chi^2$ follows
the definition in Equation \ref{eq:chi2}. 
We do not have a closed analytic expression for $\hat{f}$, we compute it
by numerical integration of the analytical approximations found in
\citet{GaravitoCamargo2014}.

@@

The low p-values are illustrated by the results shown in Figure \ref{multiphaseresults}.
Left column shows the difference between the integrated distributions
of the full sample (2500 input models) and the $1\%$ models with the lowest
$\chi^2/d.o.f.$, where we use the total number of degrees of freedom,
$d.o.f.=104$. 
The right column shows the actual $\chi^2/d.o.f.$ 
and its corresponding parameter value. 
Under these conditions we find $\sigma_{\rm cl}=$\sigmaclump,
$v_{\infty {\rm, cl}}=$\inftyclump\ and $P_{\rm cl}=$\probaclump, {\bf
  with the minimum $\chi^2/d.o.f.=11.7$.}

Qualitatively this result can be understood as follows.
Due to the large fraction of \lya photons being emitted within the
moving clumps ($P_{\rm cl}\sim 1$) the `intrinsic' profile follows 
the clump kinematics. In other words, the width of the intrinsic line
is set by $\sigma_{\mathrm{cl}}$ and its median offset mainly by
$v_{\infty,{\rm cl}}$ and $\beta_{\rm cl}$, {\bf where $\beta_{\rm cl}$ is
the exponent defining the steepness of the radial velocity profile. }
Furthermore, the relatively low mean number of clumps per line of
sight, $0.8<f_{\mathrm{cl}}<8$, combined with the high velocity dispersion
of the clumps implies existence of low-density inter-clump 
regions where \lya photons can freely propagate.
This gives result to an emergent spectrum close to the intrinsic one, explaining the high
flux at the line's center.

@@

Following the methodology of \citet{2011ApJ...726..108T} we estimate
that a dark matter halo with a virial mass $\sim 6\times 10^{11}$
M$_{\odot}$ {\bf and a virial radius $220$ kpc } could explain this dynamical mass.
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the median mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P,2016ApJ...832..169P}.  
{\bf This estimate should be considered an upper bound, lower values
  could be achieved if one considers instead a cored profile.}
Figure \ref{fig:mass} shows the enclosed mass as a function of radius
for different dark matter halos together with the \tol's dynamical
mass estimate. 
{\bf This estimate and the validity of its assumptions can only be
  verified by detailed mass modelling once \tol's full rotation curve
  becomes available.}

{\bf New IFU observations are needed to confirm \tol's kinematics.
This could be done with the Multi Unit Spectroscopic Explorer (MUSE) 
\citep{2014Msngr.157...13B}, the Gemini Multi-Object Spectrographs (GMOS) 
\citep{2004PASP..116..425H} or the Fibre Large Array Multi Element
Spectrograph (FLAMES) \citep{2002Msngr.110....1P} as they have the spatial
resolution and field of view required to perform the kind of study
presented by \citet{2015A&A...577A..21C} on CDGs or \citet{Herenz16}
on nearby \lya\ emitting galaxies.}

\begin{figure}
\begin{center}
\includegraphics[width=0.46\textwidth]{enclosed_mass.pdf}
\caption{Enclosed mass in a spherical system as a function of radius.
  Lines correspond to the expectation for dark matter halos following
  an NFW profile with
  different virial masses as shown in the legend. The circle corresponds
  to the dynamical mass estimates for \tol.  
  Under this conditions the dynamical mass estimates for \tol\ are consistent with the
  galaxy being hosted by a dark matter halo of
  $\sim 6\times10^{11}$ M$_{\odot}$ in mass.
    \label{fig:mass}} 
\end{center}
\end{figure}
